Formal Modeling and Verification of a Rate-Monotonic Scheduling Implementation with Real-Time Maude

### 2015-11-07 RMS/submission.tex
polished a bit wrt remarks from Prof. Song.

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEtitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\IEEEtitleabstractindextext{%
\begin{abstract}
Rate-Monotonic Scheduling (RMS) is one of the most important real-time
scheduling algorithms. There are large numbers of results about RMS,
especially on its schedulability. However, these theoretical results
contain not enough details to be used directly to investigate a
realistic RMS implementation. On the other hand, the correctness of
such an implementation is of the same importance. In this paper, we
analyze a realistic RMS implementation by formal modeling using
Real-Time Maude, a language and tool based on rewriting
logic. Overhead and some details of the hardware are taken into
account in the model. We verify the schedulability and the correctness
of the implementation within several scenarios, using model checking
tool built in Real-Time Maude. The soundness and the completeness of
our results are proved in the end.
\end{abstract}

@@

% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps (small caps for compsoc).
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.
\IEEEPARstart{P}{eriodic} task scheduling is one of the most important
topics within the field of real-time systems, due to the large number
of control systems that require cyclic
activities~\cite{buttazzo2011hard}. A set of periodic tasks is said to
be \emph{schedulable} with respect to some scheduling algorithm if all
jobs meet their deadlines. \emph{Rate-Monotonic Scheduling}
(\emph{RMS}) is a \emph{fixed} priority scheduling algorithm for
preemptive hard real-time environments proposed by Liu and
Layland~\cite{DBLP:journals/jacm/LiuL73}, which assigns priorities to
jobs according to the periods of the corresponding tasks: the smaller
period, the higher priority. RMS is proved to be the \emph{optimal}
fixed priority scheduling algorithm~\cite{DBLP:journals/jacm/LiuL73},
in the sense that any set of tasks, which is schedulable under
\emph{some} fixed priority scheduling algorithm, is also schedulable
with respect to RMS. It is widely used in safety-critical real-time
applications, such as vehicles and avionics, thanks to its optimality
and easiness to implement.

@@

\end{picture}
\caption{An Counterexample of Scenario (iv).}
\label{f:counterexample}
\end{figure*}
}

@@

\end{tikzpicture}
\caption{An counterexample of schedulablity in Scenario~(iv). Each job
  of each task executes for the maximum of its computation
  requirement. The first job of $\tau_3$ misses its deadline at time
  $15ms$, which is the initiation time for the second job of
  $\tau_3$.}
\label{f:counterexample}
\end{figure*}

@@

\section{Conclusion}
\label{s:conclusion}
We have formally modeled a realistic implementation of RMS using
Real-Time Maude, a modeling language based on rewriting logic. By
taking into acount the overhead of scheduling and switching, assuming
that the computation requirement for each task may not be constant,
and modeling some mechanism of the hardware, we believe that our model
contains enough details to be analyzed for the behaviors of the real
system. Two important properties--schedulability and correctness--are
verified by model checking on our model, within several given
scenarios. We also prove at last the soundness and completeness of our
results.

>>>>

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEtitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\IEEEtitleabstractindextext{%
\begin{abstract}
Rate-Monotonic Scheduling (RMS) is one of the most important real-time
scheduling used in industry. There are a large number of results about
RMS, especially on its schedulability. However, the theoretical
results do not contain enough details to be used directly for an
industrial RMS implementation. On the other hand, the correctness of
such an implementation is of the crucial importance. In this paper, we
analyze a realistic RMS implementation by using Real-Time Maude, a
formal modeling language and analysis tool based on rewriting
logic. Overhead and some details of the hardware are taken into
account in the model. We validate the schedulability and the
correctness of the implementation within key scenarios. The soundness
and the completeness of our approach are substantiated.
\end{abstract}

@@

% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps (small caps for compsoc).
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.
\IEEEPARstart{P}{eriodic} task scheduling is one of the most important
topics within the field of industrial real-time systems.
%due to the large number of control systems that require cyclic activities~\cite{buttazzo2011hard}. 
A set of periodic tasks is said to be \emph{schedulable} with respect
to some scheduling algorithm if all jobs meet their
deadlines. \emph{Rate-Monotonic Scheduling} (\emph{RMS}) is a
\emph{fixed} priority scheduling algorithm for preemptive hard
real-time environments proposed by Liu and
Layland~\cite{DBLP:journals/jacm/LiuL73}, which assigns priorities to
jobs according to the periods of the corresponding tasks: the smaller
period, the higher priority. RMS is proved to be the \emph{optimal}
fixed priority scheduling algorithm~\cite{DBLP:journals/jacm/LiuL73},
in the sense that any set of tasks, which is schedulable under
\emph{some} fixed priority scheduling algorithm, is also schedulable
with respect to RMS. It is widely used in safety-critical real-time
applications, such as vehicles and avionics, thanks to its optimality
and easiness to implement.

@@

\end{picture}
\caption{A Counterexample of Scenario (iv).}
\label{f:counterexample}
\end{figure*}
}

@@

\end{tikzpicture}
\caption{A counterexample of schedulablity in Scenario~(iv). Each job
  of each task executes for the maximum of its computation
  requirement. The first job of $\tau_3$ misses its deadline at time
  $15ms$, which is the initiation time for the second job of
  $\tau_3$.}
\label{f:counterexample}
\end{figure*}

@@

\section{Conclusion}
\label{s:conclusion}
We have formally modeled a realistic implementation of RMS using
Real-Time Maude, a modeling language based on rewriting logic. By
taking into acount the overhead of scheduling and switching, assuming
that the computation requirement for each task may not be constant,
and modeling some mechanism of the hardware, our model contains
sufficient details to be analyzed for the behaviors of real
systems. Two important properties--schedulability and correctness--are
verified by model checking on our model, within several key
scenarios. We demonstrate the soundness and completeness of our
results.
